---
output:
  html_document: default
  author: Jared White (script)
  date: 2024-03-20
  license: cc-by-4.0
---

```{r message=FALSE, warning=FALSE, include=FALSE, tidy=TRUE}
# Load libraries, import, clean, and format the data
library(tidyverse)
library(janitor)
library(readr)
library(lubridate)
library(formatR)

df_raw <- clean_names(read_csv("li_data.csv"))
# convert dates to iso & engagement to numeric, remove "#" from strings.
df <- df_raw %>% mutate(date = mdy(date),
                              engagement = as.numeric(sub("%", "", engagement))) %>% 
  mutate(across(starts_with("hashtag"), ~str_remove(., "#")))
remove(df_raw)
# make long for hashtag analysis
df_long <- df %>% 
  pivot_longer(cols = starts_with("hashtag"),
              names_to = "hashtag_number",
              values_to = "hashtag") %>% drop_na() %>% 
  select(-hashtag_number)

# make useful variables
count_posts_total <- max(df$id)
rpt_start_date <- format(min(df$date), "%b %d, '%y")
rpt_end_date <- format(max(df$date), "%b %d, '%y")
cap_text <- paste0(count_posts_total, " posts: ", 
                   rpt_start_date," - ",rpt_end_date)

```

## LinkedIn Analytics Report

##### `r paste0(cap_text)`

##### The engagement formula is:

$\large Engagement =\huge \frac{{n_{reactions} + (n_{comments} \times 2)}}{n_{views}}$

(Engagement is rated as a percentage of total views, and comments are weighted twice as heavily as reactions.)

#### Subject - Engagement

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
sub_eng <- df %>% group_by(subject) %>% summarize(avg_eng = mean(engagement))

#Plot
ggplot(data = sub_eng, mapping = aes(x=subject,y=avg_eng,fill=subject)) + geom_col() + 
  labs (y = "Average Engagement", x = "Subject of Post", title = "Average Engament per Subject", caption = cap_text) + 
  geom_text(aes(x = subject, y = 0.8*avg_eng, label = paste0(round(avg_eng, 2), "%"), fontface = 'bold'))
```

#### Engagement Time-Series

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
ggplot(data = df, aes(x = date, y = engagement)) + 
  geom_smooth(color = 'black')+ 
    geom_text(aes(label = id, color = subject, fontface = 'bold')) +
  labs(title = "Engagement per Post",
       subtitle = "(shown over time)",
       x = "Date of Post",
       y = "Engagement Metric to Date",
       caption = cap_text)
```

#### Engagement 5 Number Summary

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
eng_5num <- fivenum(df$engagement, na.rm = TRUE) 
names(eng_5num) <- c("Minimum", "Q1", "Median", "Q3", "Maximum")
print(eng_5num)
```

(Engagement metric for all posts reported)

#### Views Time-Series

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
ggplot(data = df, aes(x = date, y = views)) + 
  geom_smooth(color = 'black')+ 
    geom_text(aes(label = id, color = subject, fontface = 'bold')) +
  labs(title = "Views per Post",
       subtitle = "(shown over time)",
       x = "Date of Post",
       y = "Number of Views to Date",
       caption = cap_text)
```

#### Views 5 Number Summary

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
views_5num <- fivenum(df$views, na.rm = TRUE) 
names(views_5num) <- c("Minimum", "Q1", "Median", "Q3", "Maximum")
print(views_5num)
```

(Views for all posts reported)

#### Hashtags - Average Views

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
# make a summary of hashtag metrics
hash_summ <- df_long %>%
  group_by(hashtag) %>% 
  summarise(n_posts = n_distinct(id),
            sum_views = sum(views),
            avg_views = (sum(views)/n_distinct(id)),
            .groups = "drop") %>% arrange(., -avg_views)
# get top tags performing used in more than one post
top_hash <- head(filter(hash_summ, n_posts > 1), n = 10L)
ggplot(data = top_hash) + 
  geom_col(aes(x = hashtag, y = avg_views, fill = hashtag), show.legend = FALSE) +
  geom_text(aes(x = hashtag, y = 0.85*avg_views, label = n_posts, fontface = "bold")) +
  guides(x = guide_axis(angle = 45)) +
  labs(title = "Top 10 Performing Hashtags, with Number of Times Used",
       subtitle = "(of tags used more than once)",
       x = "Hashtag",
       y = "Average Views to date",
       caption = cap_text)
```

#### Post Day - Views

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
day_view <- df %>% group_by(wday) %>% summarize(avg_views = mean(views))
#add day numbers to order by (FIX THIS)
for (i in 1:length(day_view$wday)){
  if (day_view$wday[i] == 'monday'){
    day_view$day_num[i] <- 1
  } else if(day_view$wday[i] == 'tuesday'){
    day_view$day_num[i] <- 2
  } else if(day_view$wday[i] == 'wednesday'){
    day_view$day_num[i] <- 3
  } else if(day_view$wday[i] == 'thursday'){
    day_view$day_num[i] <- 4
  } else if(day_view$wday[i] == 'friday'){
    day_view$day_num[i] <- 5
  } else if(day_view$wday[i] == 'saturday'){
    day_view$day_num[i] <- 6
  } else if(day_view$wday[i] == 'sunday'){
    day_view$day_num[i] <- 7
  }}
#Plot
ggplot(data = day_view, 
       mapping = aes(x=factor(wday, levels = wday[order(day_num)]),y=avg_views)) + 
  scale_fill_distiller(palette = "Blues", name = NULL, guide = NULL) +
  geom_col(aes(fill = day_num),color = 'darkblue') +
  geom_text(aes(x = wday, y = 0.8*avg_views, label = wday, fontface = "bold")) +
  theme(axis.text.x = element_blank())+
  labs(title = "Average Views per Weekday of Post",
       x = "Day of Week",
       y = "Average Number of Views",
       caption = cap_text)
```


#### Image - Engagement

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
img_eng <- df %>% group_by(image) %>% summarize(avg_eng = mean(engagement))
#Plot
ggplot(data = img_eng, mapping = aes(x=image,y=avg_eng,fill = factor(image))) + 
  geom_col(color = "darkblue") +
  scale_fill_brewer(palette = "Blues", name = NULL, guide = NULL) +
  geom_text(aes(x = image, y = 0.85*avg_eng, 
                label = ifelse(image == TRUE, paste0(round(avg_eng, 2), "%", "\n", "Image"),
                               paste0(round(avg_eng, 2), "%", "\n", "No Image")),
                fontface = 'bold')) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Average Post Engagement With & Without Images",
       x = "Image Included", y = "Average Engagement", caption = cap_text)
```

#### Link - Engagement

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
link_eng <- df %>% group_by(link) %>% summarize(avg_eng = mean(engagement))
#Plot
ggplot(data = link_eng, mapping = aes(x=link,y=avg_eng,fill = factor(link))) + 
  geom_col(color = "darkgreen") +
  scale_fill_brewer(palette = "Greens", name = NULL, guide = NULL) +
  geom_text(aes(x = link, y = 0.85*avg_eng, 
                label = ifelse(link == TRUE, paste0(round(avg_eng, 2), "%", "\n", "Link"),
                               paste0(round(avg_eng, 2), "%", "\n", "No Link")),
                fontface = 'bold')) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Average Post Engagement With & Without Links",
       x = "Link Included", y = "Average Engagement", caption = cap_text)
```

#### Emojis - Engagement

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
emoji_eng <- df %>% group_by(multiple_emojis) %>% summarize(avg_eng = mean(engagement))
#Plot
ggplot(data = emoji_eng, mapping = aes(x=multiple_emojis,y=avg_eng,fill = factor(multiple_emojis))) + 
  geom_col(color = "darkred") +
  scale_fill_brewer(palette = "Reds", name = NULL, guide = NULL) +
  geom_text(aes(x = multiple_emojis, y = 0.8*avg_eng, 
                label = ifelse(multiple_emojis == TRUE, paste0(round(avg_eng, 2), "%", "\n", "2+ Emojis"),
                               paste0(round(avg_eng, 2), "%", "\n", "1 or No Emojis")),
                fontface = 'bold')) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Average Post Engagement With & Without Multiple Emojis",
       x = "Multiple Emojis Included", y = "Average Engagement", caption = cap_text)
```

#### Cross Posted - Engagement

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Create a summary table
cross_eng <- df %>% group_by(cross_posted) %>% summarize(avg_eng = mean(engagement))
#more colors
gold_boo <- c("#fff54e","#fffab2") 
names(gold_boo) <- c("TRUE","FALSE")
#Plot
ggplot(data = cross_eng, mapping = aes(x=cross_posted,y=avg_eng,fill = cross_posted)) + 
  geom_col(color = "gold")+
  scale_fill_manual(values = gold_boo, name = NULL, guide = NULL) +
  geom_text(aes(x = cross_posted, y = 0.8*avg_eng, 
                label = ifelse(cross_posted == TRUE, paste0(round(avg_eng, 2), "%", "\n", "Cross Posted"),
                               paste0(round(avg_eng, 2), "%", "\n", "Not Cross Posted")),
                fontface = 'bold')) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Whether the post was also Cross-Posted to a Relevant LinkedIn Group",
       x = "Cross Posted", y = "Average Engagement", caption = cap_text)
```

#### Several Metrics Scatter Plot

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
#Plot
ggplot(data = df, mapping = aes(x=wordcount,y=engagement,color=subject)) + 
  geom_point(aes(size = views)) + 
  geom_text(aes(label = id), vjust =ifelse(
    df$engagement > .9*max(df$engagement), 1.5, -.75), hjust = 0.5, color = 'black') +
   labs(title = "Wordcount to Engagement for Each Post",
       x = "Wordcount",
       y = "Post Engagement",
       caption = cap_text)
```

#### Wordcount 5 Number Summary

```{r echo=FALSE, warning=FALSE, tidy=TRUE}
wc_5num <- fivenum(df$wordcount, na.rm = TRUE) 
names(wc_5num) <- c("Minimum", "Q1", "Median", "Q3", "Maximum")
print(wc_5num)
```

#### Script Citations

R:

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

tidyverse:

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L,
Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP,
Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.”
  _Journal of Open Source Software_, *4*(43), 1686. doi: 10.21105/joss.01686 (URL:
  https://doi.org/10.21105/joss.01686).

janitor:

Sam Firke (2023). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package
  version 2.2.0. https://CRAN.R-project.org/package=janitor

readr:

Hadley Wickham, Jim Hester and Jennifer Bryan (2024). readr: Read Rectangular Text Data. R
  package version 2.1.5. https://CRAN.R-project.org/package=readr
  
lubridate:

  Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal
  of Statistical Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.


